---
title: "iris data exploration"
output:
  html_document:
    df_print: paged
---

# Group Members
## Laxman Panthi

## Zubin Shah



```{r}
#libraries
library(tidyverse)
library(corrplot)
```


# Data Pre-processing
## Load the iris data

```{r}
data(iris)
```


## Data Exploration
```{r}
head(iris)
summary(iris)
```

```{r}
glimpse(iris)
```


## Correlation Analysis
```{r}
irisNum <- iris%>%select(Petal.Length,Petal.Width,Sepal.Length,Sepal.Width)

corrplot(cor(irisNum), type = "full")
```

# Analysis
We are trying to solve it as a classification problem and see if we will be able to develop a model that can identify the classes.

## K-Means Clustering
```{r}
set.seed(20)
irisCluster <- kmeans(iris[, 1:4], centers = 3, nstart = 100)
irisCluster
```


```{r}
table(irisCluster$cluster, iris$Species)
```

```{r}
plot(iris[c("Sepal.Length", "Sepal.Width")], col=irisCluster$cluster)
points(irisCluster$centers[,c("Sepal.Length", "Sepal.Width")], col=1:3, pch=8, cex=2)
```


# Results
Using k-means clustering to cluster the data, we clustered the data into three clusters trying to mimic the cluster we already have. The k-means cluster pass 1 was able to identify the setosa class really goos (100% accuracy), the versicolor class with 96% accuracy, not so great for the virginica.



## Regression problem - in progress

```{r}
set.seed(9876)
irisGlm <- iris
irisGlm$Species <- factor(irisGlm$Species)
train <- sample(1:nrow(irisGlm), nrow(irisNum)*.7)
test <- -train

trainData <- irisGlm[train,]
testData <- irisGlm[test,]
```

```{r}
glm_model1 <- glm(Species~.,data = trainData, family = binomial(link="logit"))
summary(glm_model1)
```

### GLM is in progress



